---
title: "Targeted scRNA-seq Preprocessing Pipeline for Mycn Fragments"
author: "ccz"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Overview
This pipeline processes targeted single-cell RNA sequencing data (both STRT-seq and 10x Genomics platforms) to quantify specific *Mycn* transcript fragments. The quantified fragment counts are then integrated into a pre-existing Seurat object to classify cells based on their transcriptomic genotypes (e.g., tKO, tCtrl).

> **Note for Reproducibility:** Absolute file paths (e.g., `/data/ccz/...`) are used in this script. Users running this code on their local machines should update these paths to match their local directory structures.

---

## 1. STRT-seq Data Processing

### 1.1 Load Packages and Barcode Reference
```{r load-strt-ref}
library(dplyr)
library(purrr)
library(tidyr)
library(Seurat)
```

### 1.2 Define STRT-seq Processing Function
```{r def-strt-func}
# Load the 96-well 8bp barcode reference for STRT-seq
# This maps the 8bp barcode to specific cell wells (sc1 to sc96)
barcode_ref <- read.table('/opt/smartliu-master/96-8bp-barcode', col.names = "barcode")
barcode_ref$type <- paste0('sc', 1:96)
# Function to process raw STRT-seq targeted fragment reads
# Steps: Read data -> Extract Barcode & UMI -> Deduplicate (remove PCR duplicates) -> Count -> Map to reference
process_targeted_gene <- function(file_path, sample_prefix, fragment_name, ref_df) {
  
  message("Processing: ", basename(file_path), " -> ", fragment_name)
  
  # Check if file is empty to prevent errors
  if (file.info(file_path)$size == 0) {
    message("⚠️ Warning: File is empty, skipping -", basename(file_path))
    return(data.frame(cell_name = character(), Fragment = character(), Expression_Count = integer()))
  }
  
  # Read raw reads
  raw_data <- read.table(file_path, col.names = "V1", fill = TRUE, stringsAsFactors = FALSE, comment.char = "")
  raw_data <- raw_data[!grepl("^@", raw_data$V1), , drop = FALSE] # Remove header lines if any
  
  if (nrow(raw_data) == 0) {
    return(data.frame(cell_name = character(), Fragment = character(), Expression_Count = integer()))
  }
  
  processed_data <- raw_data %>%
   rename(cell_barcode = V1, umi = V2) %>%
    # UMI deduplication: keep only unique combinations of cell_barcode and UMI
    distinct(cell_barcode, umi, .keep_all = TRUE) %>% 
    # Count unique UMIs per cell
    count(cell_barcode, name = "Expression_Count") %>%
    # Map to 96-well reference
    left_join(ref_df, by = c("cell_barcode" = "barcode")) %>%
    filter(!is.na(type)) %>% 
    mutate(
      cell_name = paste0(sample_prefix, type), # Create unique cell ID matching Seurat object
      Fragment = fragment_name 
    ) %>%
    select(cell_name, Fragment, Expression_Count) 
  
  return(processed_data)
}
```

### 1.3 Process STRT-seq Knockout (KO) and Wildtype (WT) Samples
```{r process-strt-samples}
# ---------------------------------------------------------
# Process KO samples
# ---------------------------------------------------------
sample_info <- data.frame(
  file_path = c(
    '/data/ccz/project/project_chiw_myc/10X/intersect/ko1_ST.txt',
    '/data/ccz/project/project_chiw_myc/10X/intersect/ko1_E2_3.txt',
    '/data/ccz/project/project_chiw_myc/10X/intersect/ko1_E1.txt',
    '/data/ccz/project/project_chiw_myc/10X/intersect/ko2_ST.txt',
    '/data/ccz/project/project_chiw_myc/10X/intersect/ko2_E2_3.txt',
    '/data/ccz/project/project_chiw_myc/10X/intersect/ko2_E1.txt',
    '/data/ccz/project/project_chiw_myc/10X/intersect/ko3_ST.txt',
    '/data/ccz/project/project_chiw_myc/10X/intersect/ko3_E2_3.txt',
    '/data/ccz/project/project_chiw_myc/10X/intersect/ko3_E1.txt'
  ),
  prefix = c(
    'K1_L2_Q2472_', 'K1_L2_Q2472_', 'K1_L2_Q2472_',
    'K2_L8_M032_',  'K2_L8_M032_',  'K2_L8_M032_',
    'K3_L8_M034_',  'K3_L8_M034_',  'K3_L8_M034_'
  ),
  fragment = c(
    'Mycn_ST', 'Mycn_E2_3', 'Mycn_E1', 
    'Mycn_ST', 'Mycn_E2_3', 'Mycn_E1', 
    'Mycn_ST', 'Mycn_E2_3', 'Mycn_E1'  
  ),
  stringsAsFactors = FALSE
)

# Iterate over KO samples and combine results
long_result <- pmap_dfr(
  list(sample_info$file_path, sample_info$prefix, sample_info$fragment),
  ~ process_targeted_gene(file_path = ..1, sample_prefix = ..2, fragment_name = ..3, ref_df = barcode_ref)
)

# Convert from long to wide format (Cells as rows, Fragments as columns)
final_matrix <- long_result %>%
  pivot_wider(names_from = Fragment, values_from = Expression_Count, values_fill = 0)

# ---------------------------------------------------------
# Process WT samples (Same logic as KO)
# ---------------------------------------------------------
sample_info_wt <- data.frame(
  file_path = c(
    '/data/ccz/project/project_chiw_myc/10X/intersect/wt/wt1_ST.txt',
    '/data/ccz/project/project_chiw_myc/10X/intersect/wt/wt1_E2_3.txt',
    '/data/ccz/project/project_chiw_myc/10X/intersect/wt/wt1_E1.txt',
    '/data/ccz/project/project_chiw_myc/10X/intersect/wt/wt2_ST.txt',
    '/data/ccz/project/project_chiw_myc/10X/intersect/wt/wt2_E2_3.txt',
    '/data/ccz/project/project_chiw_myc/10X/intersect/wt/wt2_E1.txt',
    '/data/ccz/project/project_chiw_myc/10X/intersect/wt/wt3_ST.txt',
    '/data/ccz/project/project_chiw_myc/10X/intersect/wt/wt3_E2_3.txt',
    '/data/ccz/project/project_chiw_myc/10X/intersect/wt/wt3_E1.txt'
  ),
  prefix = c(
    'W1_L1_Q3029_', 'W1_L1_Q3029_', 'W1_L1_Q3029_',
    'W2_L1_Q3030_',  'W2_L1_Q3030_',  'W2_L1_Q3030_',
    'W3_L2_Q2471_',  'W3_L2_Q2471_',  'W3_L2_Q2471_'
  ),
  fragment = c(
    'Mycn_ST', 'Mycn_E2_3', 'Mycn_E1', 
    'Mycn_ST', 'Mycn_E2_3', 'Mycn_E1', 
    'Mycn_ST', 'Mycn_E2_3', 'Mycn_E1'  
  ),
  stringsAsFactors = FALSE
)

long_result_wt <- pmap_dfr(
  list(sample_info_wt$file_path, sample_info_wt$prefix, sample_info_wt$fragment),
  ~ process_targeted_gene(file_path = ..1, sample_prefix = ..2, fragment_name = ..3, ref_df = barcode_ref)
)

final_matrix_wt <- long_result_wt %>%
  pivot_wider(names_from = Fragment, values_from = Expression_Count, values_fill = 0)
```

---

## 2. 10x Genomics Data Processing

### 2.1 Define 10x Processing Function
```{r def-10x-func}
# Function to process 10x Genomics targeted fragment reads
process_10x_targeted <- function(file_path, sample_name, fragment_name) {
  message(paste("Processing:", fragment_name))
  
  # Read data (V1 = Barcode, V2 = UMI)
  df <- read.table(file_path, fill = TRUE, stringsAsFactors = FALSE, col.names = c("V1", "V2"))
  df <- df[!grepl("^@", df$V1), ]
  
  result <- df %>%
    rename(cell_barcode = V1, umi = V2) %>%
    # Clean barcode string (remove specific prefix 'Z:' generated by upstream alignment)
    mutate(cell_barcode = gsub('Z:', '', cell_barcode)) %>%
    # UMI deduplication
    distinct(cell_barcode, umi) %>%
    # Count unique UMIs per cell
    count(cell_barcode, name = "Expression_Count") %>%
    mutate(
      Sample = sample_name,
      Fragment = fragment_name
    ) %>%
    select(Sample, Fragment, cell_barcode, Expression_Count)
  
  return(result)
}
```

### 2.2 Process 10x Batches
```{r process-10x-samples}
# ---------------------------------------------------------
# Batch 1: 10x Samples
# ---------------------------------------------------------
sample_info_10x <- data.frame(
  file_path = c(
    '/data/ccz/project/project_chiw_myc/10X/intersect/Mycn_ST.txt',
    '/data/ccz/project/project_chiw_myc/10X/intersect/Mycn_E2_3.txt',
    '/data/ccz/project/project_chiw_myc/10X/intersect/Mycn_E1.txt'
  ),
  sample_name = rep("Sample1", 3), 
  fragment_name = c("Mycn_ST", "Mycn_E2_3", "Mycn_E1")
)

final_10x_result <- pmap_dfr(sample_info_10x, function(file_path, sample_name, fragment_name) {
  process_10x_targeted(file_path, sample_name, fragment_name)
})

```
## 3. Integration with Seurat Object & Transcript Genotyping

In this section, we map the quantified fragment counts back to the main Seurat object (`sub_combined`), normalize the counts, and classify cells into different transcriptomic genotypes (tKO, tCtrl, tHT, tNone) based on the presence or absence of specific *Mycn* fragments.

```{r integrate-seurat}
# Note: 'sub_combined' is assumed to be a pre-loaded Seurat object in the environment.

# 1. Initialize fragment count columns to 0 to prevent NA issues
sub_combined$Mycn_ST <- 0
sub_combined$Mycn_E2_3 <- 0
sub_combined$Mycn_E1 <- 0

# 2. Map STRT-seq KO counts to Seurat metadata
# Using intersect() ensures we only update cells that exist in the Seurat object
rownames(final_matrix) <- final_matrix$cell_name
valid_cells_ko <- intersect(final_matrix$cell_name, Cells(sub_combined))
sub_combined@meta.data[valid_cells_ko, 'Mycn_ST'] <- final_matrix[valid_cells_ko, 'Mycn_ST']
sub_combined@meta.data[valid_cells_ko, 'Mycn_E2_3'] <- final_matrix[valid_cells_ko, 'Mycn_E2_3']
sub_combined@meta.data[valid_cells_ko, 'Mycn_E1'] <- final_matrix[valid_cells_ko, 'Mycn_E1']

# 3. Map STRT-seq WT counts to Seurat metadata
rownames(final_matrix_wt) <- final_matrix_wt$cell_name
valid_cells_wt <- intersect(final_matrix_wt$cell_name, Cells(sub_combined))
sub_combined@meta.data[valid_cells_wt, 'Mycn_ST'] <- final_matrix_wt[valid_cells_wt, 'Mycn_ST']
sub_combined@meta.data[valid_cells_wt, 'Mycn_E2_3'] <- final_matrix_wt[valid_cells_wt, 'Mycn_E2_3']
sub_combined@meta.data[valid_cells_wt, 'Mycn_E1'] <- final_matrix_wt[valid_cells_wt, 'Mycn_E1']

# 4. Format and map 10x Batch 1 counts
wide_data_10x <- final_10x_result %>%
  select(cell_barcode, Fragment, Expression_Count) %>%
  pivot_wider(names_from = Fragment, values_from = Expression_Count, values_fill = list(Expression_Count = 0)) %>% 
  as.data.frame()
rownames(wide_data_10x) <- wide_data_10x$cell_barcode

valid_cells_10x_1 <- intersect(wide_data_10x$cell_barcode, Cells(sub_combined))
sub_combined@meta.data[valid_cells_10x_1, 'Mycn_ST'] <- wide_data_10x[valid_cells_10x_1, "Mycn_ST"]
sub_combined@meta.data[valid_cells_10x_1, 'Mycn_E2_3'] <- wide_data_10x[valid_cells_10x_1, "Mycn_E2_3"]
sub_combined@meta.data[valid_cells_10x_1, 'Mycn_E1'] <- wide_data_10x[valid_cells_10x_1, "Mycn_E1"]

# 5. Format and map 10x Batch 2 counts
wide_data_2_10x <- final_2_10x_result %>%
  select(cell_barcode, Fragment, Expression_Count) %>%
  pivot_wider(names_from = Fragment, values_from = Expression_Count, values_fill = list(Expression_Count = 0)) %>% 
  as.data.frame()
rownames(wide_data_2_10x) <- wide_data_2_10x$cell_barcode

valid_cells_10x_2 <- intersect(wide_data_2_10x$cell_barcode, Cells(sub_combined))
sub_combined@meta.data[valid_cells_10x_2, 'Mycn_ST'] <- wide_data_2_10x[valid_cells_10x_2, "Mycn_ST"]
sub_combined@meta.data[valid_cells_10x_2, 'Mycn_E2_3'] <- wide_data_2_10x[valid_cells_10x_2, "Mycn_E2_3"]
sub_combined@meta.data[valid_cells_10x_2, 'Mycn_E1'] <- wide_data_2_10x[valid_cells_10x_2, "Mycn_E1"]

# ---------------------------------------------------------
# Normalization and Genotype Classification
# ---------------------------------------------------------

# Normalize fragment counts using standard log-normalization (CP100k: count per 100,000)
# Formula: log1p( (fragment_count / total_cellular_RNA_count) * 100,000 )
sub_combined$Mycn_ST  <- log1p((sub_combined$Mycn_ST / sub_combined$nCount_RNA) * 1e5)
sub_combined$Mycn_E23 <- log1p((sub_combined$Mycn_E2_3 / sub_combined$nCount_RNA) * 1e5)
sub_combined$Mycn_E1  <- log1p((sub_combined$Mycn_E1 / sub_combined$nCount_RNA) * 1e5)

# Define transcriptomic genotype (tKO) based on the presence of specific fragments
# tNone: Neither fragment is detected
# tCtrl: Control fragment (E23) detected, targeted fragment (ST) missing
# tKO:   Targeted fragment (ST) detected, control fragment (E23) missing
# tHT:   Both fragments detected (Heterozygous/Mixed)
sub_combined$pick <- ifelse(sub_combined$Mycn_ST == 0 & sub_combined$Mycn_E23 == 0, 'tNone',
                     ifelse(sub_combined$Mycn_ST == 0 & sub_combined$Mycn_E23 > 0,  'tCtrl',
                     ifelse(sub_combined$Mycn_ST > 0  & sub_combined$Mycn_E23 == 0, 'tKO', 
                            'tHT')))

# Create a combined grouping variable (Original Group + Transcript Genotype)
sub_combined$g_tGroup <- paste(sub_combined$gGroup, sub_combined$pick, sep = '_')
```


