---
title: "Single-cell data processing workflow"
author: "ccz"
date: "2026-06-09"
output: html_document
---

```{r package}
#R 4.1.3
library(Seurat)
library(dplyr)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(SeuratExtend)
library(monocle)
library(clusterProfiler)
library(org.Mm.eg.db)
library(GSEABase)
library(rio)
library(DoubletFinder)
library(JASPAR2024)
library(motifmatchr)
library(TFBSTools)
library(EnsDb.Mmusculus.v79)
library(BSgenome.Mmusculus.UCSC.mm10)
library(GenomicRanges)
library(xgboost)
library(hdWGCNA)
```
#QC
##1_STRT_Seq
```{r}
Mycn <- fread("/data/ccz/project/project_chiw_myc/out_smart/outs/refgene.refgenev3.umi.txt")#data.table
EGCC <- import("/data/Common_Data/ref/ERCC/ERCC92_RGC.gtf")
Mycn %<>%as.data.frame() 
rownames(Mycn) <- Mycn$gene_symbol
Mycn <- Mycn[,-1]
ID <- EGCC$gene_id
ERCC <- grep("^ERCC",rownames(Mycn),value = T)
mycn <-Mycn[!rownames(Mycn)%in%ERCC,] 
mycn_seu <- CreateSeuratObject(mycn)
VlnPlot(mycn_seu,features = c("nCount_RNA","nFeature_RNA"))
mycn_seu[["percent_mt"]] <- PercentageFeatureSet(mycn_seu,pattern = "^mt-")
summary(mycn$nCount_RNA,quantiles = c(0.25, 0.75))
mycn <- subset(mycn_seu,nCount_RNA > 2000&nCount_RNA<1e+06&nFeature_RNA>1000)
mycn$genotype <- mycn$orig.ident
mycn$stage <- 'E10.0'
mycn$batch <- 'STRT-seq'
```
##2_10X
```{r}
data <- Read10X('./hash_result/outs/filtered_feature_bc_matrix/')
antibody_data <- data$`Antibody Capture` %>% as.data.frame()
hto_data <- antibody_data[c('HTO1','HTO2','HTO3'),]
antibody_data <- antibody_data[c('CD201','c-Kit'),]
umi <- data$`Gene Expression`
hto_mycn <- hto_data[colSums(hto_data)>0]
join_data <- intersect(colnames(umi),colnames(hto_mycn))
umi <- umi[,join_data]
hto_mycn <- as.matrix(hto_mycn[,join_data])
mycn <- CreateSeuratObject(counts = umi)
mycn[['HTO']] <- CreateAssayObject(counts = hto_mycn)
mycn <- NormalizeData(mycn,assay = 'HTO',normalization.method = 'CLR')
mycn <- HTODemux(mycn,assay = 'HTO',positive.quantile = 0.99)
table(mycn$HTO_classification.global)
Idents(mycn) <- mycn$HTO_classification.global
mycn_data <- subset(mycn,idents = 'Singlet')
mycn_data$batch <- '2_10X'
mycn_data$stage <- "E10.5-E11"
mycn_data$genotype <- plyr::mapvalues(mycn_data$hash.ID,from = c("HTO1","HTO2","HTO3"),to=c("WT","KO1","KO2"))
mycn_ <- CreateSeuratObject(counts = GetAssayData(mycn_data,assay = 'RNA',slot = 'counts'),meta.data = mycn_data@meta.data,min.cells = 3,min.features = 200)
mycn_[['percent.mt']] <- PercentageFeatureSet(mycn_,pattern = '^mt-')
mycn_sub <- subset(mycn_,subset = nCount_RNA>5000&nFeature_RNA>1000&percent.mt<10)
mycn_sub %<>%NormalizeData() %>% FindVariableFeatures() %>% ScaleData() %>% RunPCA() %>% RunUMAP(dims = 1:20) %>% FindNeighbors(dims = 1:20) %>% FindClusters(resolution = 0.8)

#finddoubler
sweep.res.list <- DoubletFinder::paramSweep_v3(mycn_sub, PCs = 1:30, sct = FALSE)
sweep.stats <- summarizeSweep(sweep.res.list, GT = FALSE)
sweep.stats
bcmvn <- find.pK(sweep.stats)
mpK <- as.numeric(as.vector(bcmvn$pK[which.max(bcmvn$BCmetric)]))
homotypic.prop <- modelHomotypic(mycn_sub$RNA_snn_res.0.8)          
nExp_poi <- round(0.075*ncol(mycn_sub))
nExp_poi.adj <- round(nExp_poi*(1-homotypic.prop))
mycn_sub <- doubletFinder_v3(mycn_sub, PCs = 1:30, pN = 0.25, pK = mpK, nExp = nExp_poi, reuse.pANN = FALSE, sct = FALSE)
Idents(mycn_sub) <- mycn_sub$DF.classifications_0.25_0.03_181
mycn_dou <- subset(mycn_sub,idents = 'Singlet')
mycn_dou$batch <- '2_10X'
```
##3_10X
```{r}
new_data <- Read10X("/home/chenzhe/project/mycn_data/3_data/filtered_feature_bc_matrix/")
new_mycn_seu <- CreateSeuratObject(counts = new_data$`Gene Expression`)
new_mycn_seu[['HTO']] = CreateAssayObject(counts = new_data$`Antibody Capture`)
DefaultAssay(new_mycn_seu) <- 'RNA'
new_mycn_seu %<>% NormalizeData() %>% FindVariableFeatures() %>% ScaleData() 
new_mycn_seu <- NormalizeData(new_mycn_seu, assay = "HTO", normalization.method = "CLR")
new_mycn_seu <- HTODemux(new_mycn_seu, assay = "HTO", positive.quantile = 0.99)
new_mycn_seu_sub <- subset(new_mycn_seu,idents=c("Negative","Doublet"),invert=TRUE)
new_mycn_seu_sub %<>% NormalizeData() %>% FindVariableFeatures() %>% ScaleData() %>% RunPCA() %>% RunUMAP(dims = 1:20) 
new_mycn_seu_sub[["percent.mt"]] <- PercentageFeatureSet(new_mycn_seu_sub, pattern = "^mt-")

VlnPlot2(new_mycn_seu_sub,features = c("nCount_RNA","nFeature_RNA","percent.mt"),group.by = 'hash.ID')
DimPlot(new_mycn_seu_sub,group.by = "hash.ID")
FeaturePlot(new_mycn_seu_sub,features = c('Cdh5','Gja5','Runx1','Kitl'))
new_sample <- CreateSeuratObject(counts = new_mycn_seu_sub@assays$RNA@counts,
                                   meta.data = new_mycn_seu_sub@meta.data,
                                   min.cells = 3,
                                   min.features = 200)
new_sample %<>% NormalizeData() 
new_sample[["percent.mt"]] <- PercentageFeatureSet(new_sample, pattern = "^mt-")
VlnPlot2(new_sample,features = c("nCount_RNA","nFeature_RNA","percent.mt"),group.by='hash.ID')
new_sample <- subset(new_sample,nCount_RNA<75000&nFeature_RNA>500&percent.mt<5)
new_sample %<>% NormalizeData() %>% FindVariableFeatures() %>% ScaleData() %>% RunPCA() %>% RunUMAP(dims = 1:30) %>% FindNeighbors(dims = 1:30) %>% FindClusters(resolution = 1)
sweep.res <- paramSweep_v3(new_sample, PCs = 1:20)
sweep.stats_object <- summarizeSweep(sweep.res, GT = FALSE)
bcmvn <- find.pK(sweep.stats_object)
pK_bcmvn <- bcmvn$pK[which.max(bcmvn$BCmetric)] %>% as.numeric()
pK_bcmvn <- as.numeric(bcmvn$pK[which.max(bcmvn$BCmetric)])
annotations <- new_sample@meta.data$RNA_snn_res.1
homotypic.prop <- modelHomotypic(annotations)
nExp_poi <- round(0.07 * nrow(new_sample@meta.data))
nExp_poi.adj <- round(nExp_poi * (1 - homotypic.prop))
new_sample <- doubletFinder_v3(new_sample, PCs = 1:20, pN = 0.25, pK = pK_bcmvn,
                                  nExp = nExp_poi.adj, reuse.pANN = FALSE, sct = F)
new_sample_filter <- new_sample[,new_sample$DF.classifications_0.25_17_256%in%'Singlet']
counts <- new_sample_filter@assays$RNA@counts
decontX_results <- decontX(counts)
new_sample_filter$Contamination <- decontX_results$contamination
new_sample_decont <- subset(new_sample_filter, subset = Contamination < 0.75)
new_sample_decont %<>% NormalizeData() %>% CellCycleScoring(s.features = cc.genes$s.genes %>% tolower() %>% Hmisc::capitalize(),g2m.features = cc.genes$g2m.genes %>% tolower() %>%Hmisc::capitalize()) 
list <- lapply(X =  SplitObject(new_sample_decont,split.by = 'hash.ID') , FUN = function(x) {
    x <- NormalizeData(x)
     x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 3000)
  })
features <- SelectIntegrationFeatures(object.list = list,nfeatures = 1800)
cellcycle <- cor(FetchData(new_sample_decont,vars = features),FetchData(new_sample_decont,vars =c("S.Score","G2M.Score")))
feature <- rownames(cellcycle)[rowSums(abs(cellcycle)>0.4)==0]
anchors <- FindIntegrationAnchors(object.list = list, anchor.features = feature)
new_sample_combined <- IntegrateData(anchorset = anchors)
new_sample_combined %<>% ScaleData() %>% RunPCA()
new_sample_combined$stage <- ifelse(new_sample_combined$hash.ID%in%c("E10-KO","E10-WT"),'E10','E9.5')
new_sample_combined$genotype <- ifelse(new_sample_combined$hash.ID%in%c("E10-KO","E9-KO"),'KO','WT')
new_sample_combined$batch <- '3_10X'

```
#Integrating multi-omics data for batch correction, dimensionality reduction, and clustering

```{r}
all_dat <- merge(new_sample_combined,y=list(mycn_dou,mycn))
all_dat %<>% NormalizeData() %>% CellCycleScoring(s.features = cc.genes$s.genes %>% tolower() %>% Hmisc::capitalize(),g2m.features = cc.genes$g2m.genes %>% tolower() %>%Hmisc::capitalize()) 
list <- lapply(X =  SplitObject(all_dat,split.by = 'stage') , FUN = function(x) {
    x <- NormalizeData(x)
     x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
  })
features <- SelectIntegrationFeatures(object.list = list,nfeatures = 2000)
cellcycle <- cor(FetchData(all_dat,vars = features),FetchData(all_dat,vars =c("S.Score","G2M.Score")))
feature <- rownames(cellcycle)[rowSums(abs(cellcycle)>0.4)==0]
anchors <- FindIntegrationAnchors(object.list = list, anchor.features = feature)
all_combined <- IntegrateData(anchorset = anchors)
all_combined %<>% ScaleData() %>% RunPCA()
DefaultAssay(all_combined) <- "integrated"
all_combined %<>% RunUMAP(dims = 1:40,min.dist = 0.5,n.neighbors = 15)  
all_combined %<>% FindNeighbors(dims = 1:40) %>% FindClusters(resolution = 0.2)

sub_all <- all_combined[,all_combined$integrated_snn_res.0.1%in%c('0','1','2','3','4','5','6')]##Remove the stromal population
DefaultAssay(sub_all) <- 'RNA'
sub_all %<>% NormalizeData() %>% CellCycleScoring(s.features = cc.genes$s.genes %>% tolower() %>% Hmisc::capitalize(),g2m.features = cc.genes$g2m.genes %>% tolower() %>%Hmisc::capitalize()) 
list <- lapply(X =  SplitObject(sub_all,split.by = 'batch') , FUN = function(x) {
    x <- NormalizeData(x)
     x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 3000)
  })
features <- SelectIntegrationFeatures(object.list = list,nfeatures = 1800)
cellcycle <- cor(FetchData(sub_all,vars = features),FetchData(sub_all,vars =c("S.Score","G2M.Score")))
feature <- rownames(cellcycle)[rowSums(abs(cellcycle)>0.4)==0]
anchors <- FindIntegrationAnchors(object.list = list, anchor.features = feature)
sub_combined <- IntegrateData(anchorset = anchors)
sub_combined %<>% ScaleData() %>% RunPCA()
DefaultAssay(sub_combined) <- "integrated"
sub_combined %<>% RunUMAP(dims = 1:30,min.dist = 0.6,n.neighbors =15) %>% FindNeighbors(dims = 1:40) %>% FindClusters(resolution = 0.5)
sub_combined <- FindSubCluster(sub_combined,cluster = '3',graph.name = 'integrated_snn',resolution =0.3 ,subcluster.name = 'hspc.cluster')
sub_combined$Cluster <- plyr::mapvalues(sub_combined$hspc.cluster,
                                            from = sub_combined$hspc.cluster %>% unique(),
                                            to=c('HSPC2','HEC','Mye prog.','AEC1','HSPC1','VEC','VEC','Mac','AEC2','MK','Lym prog.','VEC'))

sub_combined$Cluster <- factor(sub_combined$Cluster,levels = c('VEC','AEC2','AEC1','HEC','HSPC1','HSPC2','Lym prog.','Mye prog.','Mac','MK'))
```

##Compare the difference in HSPC1 numbers
```{r}
counts_matrix <- matrix(
  c( 636, 136, 127,   #Ctrl
     812, 199, 114), #KO
  nrow = 2,
  byrow = TRUE
)
rownames(counts_matrix) <- c("Ctrl", "KO")
colnames(counts_matrix) <- c("AEC1", "HEC", "HSPC1")

target_cell <- "HSPC1"
target_counts <- counts_matrix[, target_cell]       
total_counts <- rowSums(counts_matrix)              
other_counts <- total_counts - target_counts        

test_data <- cbind(Target = target_counts, Others = other_counts)

colnames(test_data)[1] <- target_cell 
chisq_res <- chisq.test(test_data)


```



##monocle
```{r}
library(monocle)
library(Seurat)
library(dplyr)
eht <- sub_combined[,sub_combined$celltype%in%c('HEC','HSPC1')]
eht %<>% NormalizeData()
p_data <- eht@meta.data
# 提取基因信息
f_data <- data.frame(gene_short_name  = row.names(eht),  row.names  = row.names(eht))  
head(f_data)
expr_matrix <- GetAssayData(object = eht, assay = "RNA", slot = "counts") 
pd<- new('AnnotatedDataFrame', data = p_data)
fd <- new('AnnotatedDataFrame', data = f_data)
cds <- newCellDataSet(expr_matrix,
                       phenoData = pd,
                       featureData = fd,
                       lowerDetectionLimit = 0.5,
                       expressionFamily=negbinomial.size())
cds <- estimateSizeFactors(cds)
cds <- estimateDispersions(cds)
cds <- detectGenes(cds, min_expr = 0.1) 
Idents(eht) <- eht$celltype
deg_eht <- FindAllMarkers(eht,only.pos = T,logfc.threshold = log2(1.6))
cds <- setOrderingFilter(cds, deg_eht$gene)
cds <- reduceDimension(cds, max_components = 2,reduction_method =  'DDRTree',residualModelFormulaStr = "~stage")
cds <- orderCells(cds,reverse=T)
p <- plot_cell_trajectory(cds, 
                          color_by = "Pseudotime", 
                          cell_size = 2, 
                          cell_link_size = 1)
p$layers[[2]]$mapping$shape <- aes(shape = celltype)$shape
final_plot <- p + 
  scale_shape_manual(values = c(17,16))+scale_color_viridis_c(option = "D", name = "Pseudotime",direction = -1)+
  guides(shape = guide_legend(override.aes = list(size = 3))) +
  theme_classic()

final_plot

p_celltype <- plot_cell_trajectory(cds, 
                          color_by = "celltype", 
                          cell_size = 2, 
                          cell_link_size = 1)
p_celltype$layers[[2]]$mapping$shape <- aes(shape = celltype)$shape
final_plot_celltype <- p_celltype + 
  scale_shape_manual(values = c(17,16))+scale_color_manual(values = c('#FF9800', '#E53935'))+
  guides(shape = guide_legend(override.aes = list(size = 4))) +
  theme_classic()+facet_wrap(~gGroup)

final_plot_celltype
```

##Area chart
```{r fig.width=6,fig.height=4}
library(ggplot2)
library(dplyr)
library(scales) 
plot_df <- pData(cds) 
plot_df <- plot_df %>% 
  filter(!is.infinite(Pseudotime)) %>%
  filter(!is.na(gGroup)) 
ggplot(plot_df, aes(x = Pseudotime, fill = gGroup)) +
   geom_density(position = "fill", alpha = 0.9, adjust = 1.5, color = NA) + 
  
  scale_y_continuous(labels = scales::percent, expand = c(0, 0)) + 
  scale_x_continuous(expand = c(0, 0)) + 
  scale_fill_manual(values = c('#549ffa','#f84040')) +
  labs(x = "Pseudotime", y = "Proportion of gGroup", fill = "Group") +
  theme_classic() +
  theme(
    axis.text = element_text(size = 12, color = "black"),
    axis.title = element_text(size = 14, face = "bold"),
    legend.position = "right" 
  )


```

##Visualizing single-cell knockout proportions
```{r fig.width=4,fig.height=5}
eht <- sub_combined[,c(sub_combined$celltype%in%c('AEC1','HEC','HSPC1')&sub_combined$pick%in%c('tCtrl','tHT','tKO'))]
table(eht$gGroup,eht$new_gtGroup)

plot_data <- eht@meta.data %>% 
  group_by(gGroup, pick) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  group_by(gGroup) %>%
  mutate(Percent = Count / sum(Count) * 100)
ggplot(plot_data, aes(x = gGroup, y = Percent, fill = pick)) +
  geom_bar(stat = "identity", position = "stack", width = 0.7) +
  geom_text(aes(label = paste0(Count, "\n(", round(Percent, 1), "%)")), 
            position = position_stack(vjust = 0.5), 
            size = 3.5, 
            color = "white",    
            fontface = "bold",  
            lineheight = 0.9) + 
  
  theme_classic() +
  labs(x = "gGroup", y = "Percentage (%)", fill = "Cluster") + 
  scale_fill_manual(values = c('#795ff6','#00b2b3','#860162')) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12, face = "bold", color = "black"),
    axis.text.y = element_text(size = 8, color = "black"),
    axis.title = element_text(size = 14, face = "bold")
  )
```
##HEC enrichment analysis
```{r}
Idents(sub_combined) <- sub_combined$g_tGroup
hec_data <- sub_combined[,c(sub_combined$Cluster%in%'HEC'&sub_combined$g_tGroup%in%c('gKO_tKO','gCtrl_tCtrl'))]
hec_data %<>% NormalizeData() 
Idents(hec_data) <- hec_data$g_tGroup
hec_deg <- FindMarkers(hec_data,ident.1 = 'gKO_tKO',ident.2 = 'gCtrl_tCtrl',logfc.threshold = log(1.3)) %>% dplyr::filter(p_val<0.05)
hec_deg$cluster <- ifelse(hec_deg$avg_log2FC>0,'gKO_tKO','gCtrl_tCtrl')
hec_deg$gene <- rownames(hec_deg)
hec_go <- compareCluster(geneClusters = gene ~ cluster,
                         data = hec_deg  ,
                         OrgDb = "org.Mm.eg.db",
                   keyType = "SYMBOL",
                   ont = "BP"
                   )

```
##Scoring the cell substrate adhesion and focal adhesion gene sets
```{r}
focal_adhesion <- kegg_mmu$KEGGPATHID2EXTID %>% dplyr::filter(from=='mmu04510') %>% pull(to) %>% unique()  
gene_info <- AnnotationDbi::select(org.Mm.eg.db, 
                    keys = "GO:0031589", 
                    columns = c("SYMBOL"), 
                    keytype = "GOALL")
cell_adhesion_genes <- unique(gene_info$SYMBOL)

eht_all <- sub_combined[,sub_combined$celltype%in%c('AEC1','HEC','HSPC1')]
eht_all %<>% NormalizeData() 
eht_all$gGroup_celltype <- paste(eht_all$gGroup,eht_all$celltype,sep = '_')
eht_all <- AddModuleScore(eht_all,
                      features = list(cell_adhesion_genes),
                      name = 'cell-substrate_adhesion')
eht_all <- AddModuleScore(eht_all,
                      features = list(focal_adhesion),
                      name = 'focal_adhesion')

VlnPlot2(eht_all, 
         features = c('cell.substrate_adhesion1','focal_adhesion'), 
         group.by = "celltype", 
         split.by = 'gGroup_celltype',
         stat.method = "wilcox.test", 
         hide.ns = F)

eht_tko <- eht_all[,eht_all$g_tGroup%in%c("gKO_tKO",'gCtrl_tCtrl')]
eht_tko %<>% NormalizeData() 
eht_tko$g_tGroup_celltype <- paste(eht_tko$g_tGroup,eht_tko$celltype,sep = '_')
eht_tko <- AddModuleScore(eht_tko,
                      features = list(cell_adhesion_genes),
                      name = 'cell-substrate_adhesion')
eht_tko <- AddModuleScore(eht_tko,
                          features = list(focal_adhesion),
                          name = 'focal_adhesion')

VlnPlot2(eht_all, 
         features = c('cell.substrate_adhesion1','focal_adhesion1'), 
         group.by = "g_tGroup_celltype",
         stat.method = "wilcox.test", 
         hide.ns = F,pt = FALSE)+labs(title = 'cell.substrate_adhesion1')+scale_fill_manual(values = c('#1565C0', 
  '#FF9800', 
  '#E53935','#1565C0', 
  '#FF9800', 
  '#E53935'))


VlnPlot2(eht_tko, 
         features = c('cell.substrate_adhesion1','focal_adhesion1'), 
         group.by = "g_tGroup_celltype",
         stat.method = "wilcox.test", 
         hide.ns = F,pt = FALSE)+labs(title = 'cell.substrate_adhesion1')+scale_fill_manual(values = c('#1565C0', 
  '#FF9800', 
  '#E53935','#1565C0', 
  '#FF9800', 
  '#E53935'))


```
##Cell cycle scoring
```{r}
eht_tko %<>% NormalizeData() %>% CellCycleScoring(s.features = cc.genes$s.genes %>% tolower() %>% Hmisc::capitalize(),g2m.features = cc.genes$g2m.genes %>% tolower() %>%Hmisc::capitalize())  
pdf('tko_cellcylce.pdf',width = 5,height = 5)
VlnPlot2(eht_tko[,eht_tko$celltype%in%c('HEC','HSPC1')], 
         features =  c("S.Score","G2M.Score"), 
         group.by = "celltype", 
         split.by = 'g_tGroup',
         stat.method = "wilcox.test", 
         hide.ns = F,pt=F)+scale_fill_manual(values = c(
  '#860162', '#795ff6') %>% rev())+black_border

dev.off()
```


##Transcription Factor Regulation Analysis/Mycn target gene scoring and enrichment analysis
```{r}
##Predicted target gene
wgcna_data <- sub_combined[,c(sub_combined$celltype%in%c('HEC','HSPC1')&sub_combined$g_tGroup%in%c('gCtrl_tCtrl','gKO_tKO'))]
compute_mycn_network <- function(
  seurat_obj, 
  model_params, 
  target_tf = "MYCN", 
  nfold = 5, 
  callbacks = NULL, 
  wgcna_name = NULL
) {

  if (is.null(wgcna_name)) {
    wgcna_name <- seurat_obj@misc$active_wgcna
  }
  CheckWGCNAName(seurat_obj, wgcna_name)
  
  motif_matrix <- GetMotifMatrix(seurat_obj)
  motif_df <- GetMotifs(seurat_obj)
  
  if (is.null(motif_df) || !("gene_name" %in% colnames(motif_df))) {
    stop("Motif info missing or 'gene_name' column not found.")
  }

  target_motif_ids <- motif_df %>% 
    dplyr::filter(gene_name == target_tf) %>% 
    .$motif_ID %>% unique()
  
  if (length(target_motif_ids) == 0) {
    stop(paste("No motif found for", target_tf))
  }

  datExpr <- as.matrix(GetDatExpr(seurat_obj, wgcna_name = wgcna_name))
  all_genes <- colnames(datExpr)
  potential_targets <- rownames(motif_matrix)[rowSums(motif_matrix[, target_motif_ids, drop=FALSE]) > 0]
  genes_use <- all_genes[all_genes %in% potential_targets]
  genes_use <- setdiff(genes_use, target_tf)

  if (is.null(callbacks)) {
    callbacks = list(xgboost::cb.cv.predict(save_models = TRUE))
  }

  importance_df <- data.frame()
  eval_df <- data.frame()
  
  pb <- utils::txtProgressBar(min = 0, max = length(genes_use), style = 3, width = 50, char = "=")

  for (i in seq_along(genes_use)) {
    cur_gene <- genes_use[i]
    setTxtProgressBar(pb, i)
    
    cur_tfs_motifs <- names(which(motif_matrix[cur_gene, ]))
    cur_tfs <- motif_df %>% 
      dplyr::filter(motif_ID %in% cur_tfs_motifs & gene_name %in% all_genes) %>% 
      .$gene_name %>% unique()
    
    cur_tfs <- unique(c(cur_tfs, target_tf))
    cur_tfs <- setdiff(cur_tfs, cur_gene)
    
    if (length(cur_tfs) < 2) next

    x_vars <- datExpr[, cur_tfs, drop=FALSE]
    y_var <- as.numeric(datExpr[, cur_gene])
    
    if (all(y_var == 0)) next

    tryCatch({
      xgb <- xgboost::xgb.cv(
        params = model_params, 
        data = xgboost::xgb.DMatrix(x_vars, label = y_var), 
        nrounds = 100, 
        showsd = FALSE, 
        nfold = nfold, 
        callbacks = callbacks, 
        verbose = FALSE
      )

      importance <- Reduce("+", lapply(1:nfold, function(j) {
        cur_model <- xgb$models[[j]] 
        imp <- xgboost::xgb.importance(feature_names = colnames(x_vars), model = cur_model)
        full_imp <- data.frame(Feature = colnames(x_vars)) %>%
          dplyr::left_join(imp, by = "Feature") %>%
          dplyr::mutate(dplyr::across(where(is.numeric), ~replace_na(.x, 0)))
        
        return(as.matrix(full_imp[,-1]))
      })) / nfold

      importance <- as.data.frame(importance)
      importance$tf <- colnames(x_vars)
      importance$gene <- cur_gene
      
      tf_cor <- as.numeric(cor(x_vars, y_var))
      importance$Cor <- tf_cor
      
      target_row <- subset(importance, tf == target_tf)
      importance_df <- rbind(importance_df, target_row)

      xgb_eval <- as.data.frame(xgb$evaluation_log)
      xgb_eval$variable <- cur_gene
      eval_df <- rbind(eval_df, xgb_eval)
      
    }, error = function(e) { message(paste("\nError in gene", cur_gene, ":", e$message)) })
  }
  close(pb)

  seurat_obj <- SetTFNetwork(seurat_obj, importance_df, wgcna_name = wgcna_name)
  seurat_obj <- SetTFEval(seurat_obj, eval_df, wgcna_name = wgcna_name)
  
  return(seurat_obj)
}

model_params <- list(
    objective = 'reg:squarederror',
    max_depth = 1,
    eta = 0.1,
    nthread=65,
    alpha=0.5
) 
wgcna_data <- compute_mycn_network(
  wgcna_data, 
  model_params, 
  target_tf = "Mycn", 
  nfold = 5, 
  callbacks = NULL, 
  wgcna_name = NULL
)

model_params <- list(
    objective = 'reg:squarederror',
    max_depth = 1,
    eta = 0.1,
    nthread=65,
    alpha=0.5
)

library(doParallel)
registerDoParallel(cores = 65)
wgcna_data <- ConstructTFNetwork(wgcna_data, model_params)
wgcna_data <- AssignTFRegulons(
    wgcna_data,
    strategy = "A",
    reg_thresh = 0,
    n_tfs = 10
)

tf_network <- GetTFNetwork(wgcna_data)
RegulonBarPlot(wgcna_data, selected_tf='Mycn',cutoff=0.2)

mycn_target_neg <- wgcna_data@misc$tutorial$tf_net %>% dplyr::filter(tf=='Mycn'&Gain>0&Cor< -0.3) %>% arrange(desc(Gain))
mycn_target_pos <- wgcna_data@misc$tutorial$tf_net %>% dplyr::filter(tf=='Mycn'&Gain>0&Cor> 0.3) %>% arrange(desc(Gain))


hec <- sub_combined[,sub_combined$celltype%in%c('HEC')]
hec <- NormalizeData(hec)
hec <- AddModuleScore(hec,features = list(mycn_target_pos$gene),
                            name = 'mycn_target_posotive')

hec <- AddModuleScore(hec,features = list(mycn_target_neg$gene),
                            name = 'mycn_target_negitive')




hec_tctrl_tko <- hec_hspc1[,hec_hspc1$new_gtGroup%in%c('gCtrl_tCtrl','gKO_tKO')]
hec_tctrl_tko %<>% NormalizeData() 
hec_tctrl_tko <- AddModuleScore(hec_tctrl_tko,
                            features = list(mycn_target_pos$gene),
                            name = 'mycn_target_posotive')

hec_tctrl_tko <- AddModuleScore(hec_tctrl_tko,
                            features = list(mycn_target_neg$gene),
                            name = 'mycn_target_negitive')

##plotting
VlnPlot2(hec, 
         features = 'mycn_target_posotive1', 
         group.by = "celltype", 
         split.by = 'gGroup',
         stat.method = "wilcox.test", 
         hide.ns = F,pt = F)+labs(title = 'mycn_target_posotive')+scale_fill_manual(values = c('#549ffa','#f84040'))+black_border+VlnPlot2(hec_tctrl_tko, 
         features = 'mycn_target_posotive1', 
         group.by = "celltype", 
         split.by = 'g_tGroup',
         stat.method = "wilcox.test", 
         hide.ns = F,pt = F)+labs(title = 'mycn_target_posotive')+scale_fill_manual(values = c('#795ff6','#860162'))+black_border


VlnPlot2(hec_tctrl_tko, 
         features = c('mycn_target_posotive1','mycn_target_negitive1'), 
         group.by = "new_gtGroup",
         stat.method = "wilcox.test", 
         hide.ns = F)

VlnPlot2(hec, 
         features = 'mycn_target_negitive1', 
         group.by = "celltype", 
         split.by = 'gGroup',
         stat.method = "wilcox.test", 
         hide.ns = F,pt = F)+labs(title = 'mycn_target_posotive')+scale_fill_manual(values = c('#549ffa','#f84040'))+black_border+VlnPlot2(hec_tctrl_tko, 
         features = 'mycn_target_negitive1', 
         group.by = "celltype", 
         split.by = 'g_tGroup',
         stat.method = "wilcox.test", 
         hide.ns = F,pt = F)+labs(title = 'mycn_target_posotive')+scale_fill_manual(values = c('#795ff6','#860162'))+black_border


```

##Network graph visualization
```{r}
mycn_target_go <- enrichGO(gene=c(mycn_target_neg$gene,mycn_target_pos$gene) %>% unique(),
                        keyType='SYMBOL',
                        ont = 'BP',
                        OrgD = "org.Mm.eg.db"
                        )

mycn_target_go_result <- mycn_target_go@result
mycn_target_filter <- mycn_target_go_result[c(1,9,10,11,12,19,23,29,30,71,78),]#Identification of distinct significantly enriched biological process terms

library(tidyverse)
library(tidygraph)
library(ggraph)

edges <- mycn_target_filter %>%
  select(Description, geneID) %>%
  separate_rows(geneID, sep = "/") %>%
  rename(from = Description, to = geneID)

all_nodes <- unique(c(edges$from, edges$to))

nodes_df <- data.frame(name = all_nodes) %>%
  mutate(
    is_pathway = name %in% mycn_target_filter$Description,
    node_type = ifelse(is_pathway, "Pathway", "Gene"),
    reg_status = case_when(
      is_pathway ~ "Pathway Center",
      name %in% mycn_target_pos$gene ~ "Positive (Red)",
      name %in% mycn_target_neg$gene ~ "Negative (Blue)",
      TRUE ~ "Neutral (Grey)"
    )
  )


graph_data <- tbl_graph(nodes = nodes_df, edges = edges, directed = FALSE)

library(ggraph)
library(ggplot2)

pdf('../../project/mycn_data/filter_data/fig/target_mycn.pdf',width = 8,height = 6)
seed(123)
ggraph(graph_data, layout = "fr") + 
  
  geom_edge_link(alpha = 0.2, color = "black", width = 0.5) + 
  
  
  geom_node_point(aes(color = reg_status, size = node_type)) + 
  
  
  geom_node_text(aes(label = name, filter = is_pathway), 
                 repel = TRUE, fontface = "bold", size = 3) +
  
  geom_node_text(
    aes(label = name, filter = !is_pathway), 
    repel = TRUE, 
    size = 3.5,            
    fontface = "italic",  
    color = "grey20",      
    max.overlaps = 20
  ) +

  scale_color_manual(values = c(
    "Pathway Center" = "#f9b256", 
    "Positive (Red)" = "#dd522a", 
    "Negative (Blue)" = "#0e7ca2",
    "Neutral (Grey)" = "#DDDDDD"
  )) +
  
  scale_size_manual(values = c("Pathway" = 7, "Gene" = 4)) +
  
  theme_void() + 
  theme(legend.position = "right") +
  labs(
    title = "Gene-Pathway Connection Network",
    subtitle = "Visualizing MYCN targets across functional modules",
    color = "Regulation & Role", 
    size = "Node Type"
  )
dev.off()

```

#Y15 single-cell data QC, dimensionality reduction and clustering
```{r}
setwd("/mnt/dataset/chenzhe/mmyy/y15/KYSY-19274_JD-KYFW-202403-0308-JSFW")
folders = c("out_251132A_KO", "out_251132A_KY", "out_251132A_WT", "out_251132A_WY")
sceList = lapply(folders, function(f){
  path = paste0(f, "/outs/filtered_feature_bc_matrix/")
  counts = Read10X(data.dir = path)
  obj = CreateSeuratObject(counts = counts, 
                           project = f, 
                           min.cells = 3, 
                           min.features = 200)
  return(obj)
})
names(sceList) = folders
combined <- merge(x = sceList[[1]], y = sceList[-1], 
                       add.cell.ids =folders)
                       
combined[["percent.mt"]] <- PercentageFeatureSet(combined, pattern = "^mt-")
merge_split_filter <- subset(merge_split,nFeature_RNA>1000&nCount_RNA<100000&percent.mt<10)
merge_split_filter %<>% NormalizeData() %>% FindVariableFeatures() %>%  ScaleData() %>% RunPCA() %>% RunUMAP(dims = 1:30) %>% FindNeighbors(dims = 1:30) %>% FindClusters(resolution = 0.2)
sweep.res <- paramSweep_v3(merge_split_filter, PCs = 1:20)
sweep.stats_object <- summarizeSweep(sweep.res, GT = FALSE)
bcmvn <- find.pK(sweep.stats_object)
pK_bcmvn <- bcmvn$pK[which.max(bcmvn$BCmetric)] %>% as.numeric()
pK_bcmvn <- as.numeric(bcmvn$pK[which.max(bcmvn$BCmetric)])
annotations <- merge_split_filter@meta.data$RNA_snn_res.0.2
homotypic.prop <- modelHomotypic(annotations)
nExp_poi <- round(0.07 * nrow(merge_split_filter@meta.data))
nExp_poi.adj <- round(nExp_poi * (1 - homotypic.prop))
merge_split_filter <- doubletFinder_v3(merge_split_filter, PCs = 1:20, pN = 0.25, pK = pK_bcmvn,
                                  nExp = nExp_poi.adj, reuse.pANN = FALSE, sct = F)
mycn_seu_new_filter <- merge_split_filter[,merge_split_filter$DF.classifications_0.25_3_614%in%'Singlet']
counts <- mycn_seu_new_filter@assays$RNA@counts
decontX_results <- decontX(counts)
mycn_seu_new_filter$Contamination <- mycn_seu_new_filter$contamination
y15_obj <- subset(mycn_seu_new_filter,Contamination<0.75)
y15_obj %<>% NormalizeData() %>% FindVariableFeatures() %>% RunHarmony('g_Group') %>% 
FindNeighbors(y15_obj,reduction = 'harmony',dims = 1:5) %>% FindClusters(resolution = 0.3)
y15_obj <- FindSubCluster(y15_obj,cluster = '0',subcluster.name = 'hspc_new',graph.name = 'RNA_snn',resolution = 0.1)
y15_obj$final_cluster <- plyr::mapvalues(y15_obj$hspc_new,
                                           from = y15_obj$hspc_new %>% unique(),
                                           to=c('Mast cell','Mac','HSPC2','Mye Prog.','EC','EC','Mac','MK','Ery','HSPC1','MK','Granulocyte'))

y15_obj$final_cluster <- factor(y15_obj$final_cluster,levels = c('EC','HSPC1','HSPC2','Mye Prog.','Mast cell','Granulocyte','Mac','MK','Ery'))
DotPlot(y15_obj,features = c('Pecam1','Cd44','Cdh5','Unc5b','Dll4','Gja5','Mecom','Hlf','Runx1','Ptprc','Prtn3','Fcgr3','Spi1','Kit','Cd34','Cpa3','Hdc','S100a8','S100a9','Cd86','C1qa','C1qb','Gp9','Gp5','Itga2b','Gata1','Klf1','Car2') %>% unique(),scale = F,group.by = 'final_cluster')+RotatedAxis()+scale_color_gradientn(colors = c("#D3D3D3", "#FF8C00", "#FF0000", "#8B0000"))+black_border


##Runx1HE
runx1_expr <- FetchData(y15_obj,vars = 'Runx1')
cdh5_expr <- FetchData(y15_obj,vars = 'Cdh5')
ptprc_expr <- FetchData(y15_obj,vars = 'Ptprc')
y15_obj@meta.data[rownames(runx1_expr),'runx1'] <- runx1_expr[rownames(runx1_expr),'Runx1']
y15_obj@meta.data[rownames(cdh5_expr),'cdh5'] <- cdh5_expr[rownames(cdh5_expr),'Cdh5']
y15_obj@meta.data[rownames(ptprc_expr),'Ptprc'] <- ptprc_expr[rownames(ptprc_expr),'Ptprc']
y15_obj$HE <- ifelse(y15_obj$runx1>0&y15_obj$cdh5>0&y15_obj$Ptprc==0,'HE','others')

ec_hspc1_new <- y15_obj[,y15_obj$final_cluster%in%c('EC','HSPC1')]
ec_hspc1_new <- ec_hspc1_new %>% NormalizeData() %>% FindVariableFeatures(nfeatures = 1000) %>% 
  ScaleData(vars.to.regress = c('S.Score','G2M.Score')) %>%  
  RunPCA() #%>% RunHarmony()

ec_hspc1_new %<>%  RunUMAP(reduction = 'pca',dims = 1:3, min.dist = 0.5)
pdf('../mycn_new/sub_umap.pdf',width = 10,height = 4)
DimPlot(ec_hspc1_new,reduction = 'umap',group.by = 'final_cluster',cols = c("#42a4f5","#e43935"),pt.size = 1)+black_border|DimPlot(ec_hspc1_new,reduction = 'umap',group.by = 'HE',cols = c('red','grey'),pt.size = 1)+black_border
dev.off()



```
##GO enrichment analysis
```{r}
diff_enrich <- function(seurat_obj, ident1 = "KY", ident2 = "KO", 
                        logfc_thresh = log2(1.3), pval_thresh = 0.05, 
                        organism = "mmu", ont = "BP") {
  deg <- FindMarkers(seurat_obj, ident.1 = ident1, ident.2 = ident2, 
                     logfc.threshold = logfc_thresh) %>%
    filter(p_val < pval_thresh) %>%
    mutate(cluster = ifelse(avg_log2FC > 0, ident1, ident2),
           gene = rownames(.))
  deg_counts <- deg %>%
    count(cluster) %>%
    mutate(cluster = factor(cluster, levels = c(ident2, ident1)))  
  
  deg_plot <- ggplot(deg_counts, aes(x = cluster, y = n, fill = cluster)) +
    geom_col(color = "black", width = 0.7) +
    scale_fill_manual(values = c("KO" = "#E31A1C", "KY" = "#1F78B4")) + 
    labs(title = "Number of differentially expressed genes", x = "Group", y = "gene counts") +
    theme_minimal() +
    theme(legend.position = "none",
          plot.title = element_text(hjust = 0.5))
  
  # GO enrichment
  go_res <- compareCluster(gene ~ cluster, data = deg, OrgDb = "org.Mm.eg.db",
                           keyType = "SYMBOL", ont = ont)
  return(list(DEG = deg, GO = go_res,DEG_plot = deg_plot))
}

library(clusterProfiler)
he <- y15_new[,y15_new$Cluster%in%'EC'&y15_new$HE%in%'HE']
he %<>%NormalizeData() 
Idents(he) <- he$g_Group
ky_wy_results <- diff_enrich(he, ident1 = "KY", ident2 = "WY", 
                        logfc_thresh =  log2(1.2), pval_thresh = 0.05, 
                        organism = "mmu", ont = "BP")

ko_wt_results <- diff_enrich(he, ident1 = "KO", ident2 = "WT", 
                        logfc_thresh = log2(1.3), pval_thresh = 0.05, 
                        organism = "mmu", ont = "BP")

ky_wt_results <- diff_enrich(he, ident1 = "KY", ident2 = "WT", 
                        logfc_thresh = log2(1.3), pval_thresh = 0.05, 
                        organism = "mmu", ont = "BP")
ky_ko_results <- diff_enrich(he, ident1 = "KY", ident2 = "KO", 
                        logfc_thresh = 0.25, pval_thresh = 0.05, 
                        organism = "mmu", ont = "BP")

wy_wt_results <- diff_enrich(he, ident1 = "WY", ident2 = "WT", 
                        logfc_thresh = log2(1.2), pval_thresh = 0.05, 
                        organism = "mmu", ont = "BP")

```
##Cell cycle scoring
```{r}

he %<>% NormalizeData() %>%  CellCycleScoring(s.features = cc.genes$s.genes %>% tolower() %>% Hmisc::capitalize(),g2m.features = cc.genes$g2m.genes %>% tolower() %>%Hmisc::capitalize())
VlnPlot2(he, 
         features =  c("S.Score","G2M.Score"), 
         group.by   = "g_Group",stat.method="wilcox.test",
         hide.ns = F,pt=F)+scale_fill_manual(values = c('#549ffa','#459298','#f84040','#ecafcb'))+black_border

hspc1 %<>% NormalizeData() %>%  CellCycleScoring(s.features = cc.genes$s.genes %>% tolower() %>% Hmisc::capitalize(),g2m.features = cc.genes$g2m.genes %>% tolower() %>%Hmisc::capitalize())
VlnPlot2(hspc1, 
         features =  c("S.Score","G2M.Score"), 
         group.by   = "g_Group",stat.method="wilcox.test",
         hide.ns = F,pt=F)+scale_fill_manual(values = c('#549ffa','#459298','#f84040','#ecafcb'))+black_border

```
##GSEA analysis and its visualization
```{r}
library(org.Mm.eg.db)
library(clusterProfiler)
library(dplyr)
go_genes <- AnnotationDbi::select(org.Mm.eg.db, 
                                  keys = "GO:0031589", 
                                  keytype = "GOALL", 
                                  columns = "SYMBOL") %>% 
            pull(SYMBOL) %>% unique()


kegg_genes <- kegg_mmu$KEGGPATHID2EXTID %>% 
              filter(from == "mmu04510") %>% 
              pull(to)

custom_gs <- data.frame(
  term = c(rep("cell-substrate adhesion", length(go_genes)), 
           rep("Focal_Adhesion", length(kegg_genes))),
  gene = c(go_genes, kegg_genes)
)


comparisons <- list(
  c("KY", "WY"), c("KO", "WT"), c("KY", "WT"), c("KY", "KO"), c("WY", "WT")
)


specific_gsea_results <- list()

for (comp in comparisons) {
  comp_name <- paste0(comp[1], "_vs_", comp[2])
  
  
  deg <- FindMarkers(he, ident.1 = comp[1], ident.2 = comp[2], logfc.threshold = 0, min.pct = 0)
  gene_list <- sort(setNames(deg$avg_log2FC, rownames(deg)), decreasing = TRUE)
  
  
  res <- GSEA(geneList = gene_list, 
              TERM2GENE = custom_gs, 
              pvalueCutoff = 1, 
              minGSSize = 1, 
              maxGSSize = 1000)
  
  specific_gsea_results[[comp_name]] <- res
}


enrichplot::gseaplot2(
 specific_gsea_results$KO_vs_WT,
  geneSetID = 'Focal_Adhesion', 
  title = "KO vs WT(NES = 1.495801)",
  color = "red",
  pvalue_table = TRUE
)|enrichplot::gseaplot2(
 specific_gsea_results$KO_vs_WT,
  geneSetID = 'cell-substrate adhesion', 
  title = "KO vs WT(NES = 1.385164)",
  color = "red",
  pvalue_table = TRUE
)
```


```{r}
he <- y15_obj[,y15_obj$celltype]

table(y15_obj$HE,y15_obj$celltype)
y15_obj$HE
he <- y15_obj[,c(y15_obj$celltype%in%'EC'&y15_obj$HE%in%'HE')]


Idents(he) <- he$g_Group
library(org.Mm.eg.db)
library(clusterProfiler)
library(dplyr)
go_genes <- AnnotationDbi::select(org.Mm.eg.db, 
                                  keys = "GO:0042254", 
                                  keytype = "GOALL", 
                                  columns = "SYMBOL") %>% 
            pull(SYMBOL) %>% unique()


custom_gs <- data.frame(
  term = c(rep("ribosome biogenesis", length(go_genes))),
  gene = c(go_genes)
)


comparisons <- list(
  c("KY", "WY"), c("KO", "WT"), c("KY", "WT"), c("KY", "KO"), c("WY", "WT")
)


specific_gsea_results <- list()

for (comp in comparisons) {
  comp_name <- paste0(comp[1], "_vs_", comp[2])
  
  
  deg <- FindMarkers(he, ident.1 = comp[1], ident.2 = comp[2], logfc.threshold = 0, min.pct = 0)
  gene_list <- sort(setNames(deg$avg_log2FC, rownames(deg)), decreasing = TRUE)
  
  
  res <- GSEA(geneList = gene_list, 
              TERM2GENE = custom_gs, 
              pvalueCutoff = 1, 
              minGSSize = 1, 
              maxGSSize = 1000)
  
  specific_gsea_results[[comp_name]] <- res
}

specific_gsea_results$KY_vs_KO@result
```


```{r fig.width=4.5,fig.height=10}
p1 <- enrichplot::gseaplot2(
  specific_gsea_results$KY_vs_WT,
  geneSetID = 'ribosome biogenesis', 
  title = "cKO_Y15 vs Ctrl(NES =-2.054381)",
  color = "red",
  pvalue_table = TRUE
)

p2 <- enrichplot::gseaplot2(
  specific_gsea_results$KO_vs_WT,
  geneSetID = 'ribosome biogenesis', 
  title = "cKO vs Ctrl(NES =-2.016914)",
  color = "red",
  pvalue_table = TRUE
)

p3 <- enrichplot::gseaplot2(
  specific_gsea_results$KY_vs_KO,
  geneSetID = 'ribosome biogenesis', 
  title = "cKO_Y15 vs cKO(NES =1.09996)",
  color = "red",
  pvalue_table = TRUE
)

library(cowplot)
pdf('/data/ccz/project/mycn_seu/revise_data/revised_fig/gsea_ribosome.pdf',width = 4.5,height = 10)
plot_grid(p1,p2, p3, ncol = 1)
dev.off()


```
```{r  fig.width=5,fig.height=10}

p1 <- enrichplot::gseaplot2(
  specific_gsea_results$KY_vs_WT,
  geneSetID = 'ribosome biogenesis', 
  title = "cKO_Y15 vs Ctrl(NES =-2.054381)",
  color = "red",
  subplots = 1:2,
  pvalue_table = TRUE
)

p2 <- enrichplot::gseaplot2(
  specific_gsea_results$KO_vs_WT,
  geneSetID = 'ribosome biogenesis', 
  title = "cKO vs Ctrl(NES =-2.016914)",
  color = "red",
  subplots = 1:2,
  pvalue_table = TRUE
)

p3 <- enrichplot::gseaplot2(
  specific_gsea_results$KY_vs_KO,
  geneSetID = 'ribosome biogenesis', 
  title = "cKO_Y15 vs cKO(NES =1.09996)",
  color = "red",
  subplots = 1:2,
  pvalue_table = T
)


```

```{r fig.width=6,fig.height=9}
library(ggplot2)
library(patchwork)
p1 <- enrichplot::gseaplot2(
  specific_gsea_results$KY_vs_WT,
  geneSetID = 'ribosome biogenesis', 
  title = "cKO_Y15 vs Ctrl(NES =-2.054381)",
  color = "red",
  subplots = 1:2,
  pvalue_table = TRUE
)

p1[[1]] <- p1[[1]] + 
  geom_hline(yintercept = 0, linetype = "solid", color = "grey50", linewidth = 0.5)


p2 <- enrichplot::gseaplot2(
  specific_gsea_results$KO_vs_WT,
  geneSetID = 'ribosome biogenesis', 
  title = "cKO vs Ctrl(NES =-2.016914)",
  color = "red",
  subplots = 1:2,
  pvalue_table = TRUE
)

p2[[1]] <- p2[[1]] + 
  geom_hline(yintercept = 0, linetype = "solid", color = "grey50", linewidth = 0.5)


p3 <- enrichplot::gseaplot2(
  specific_gsea_results$KY_vs_KO,
  geneSetID = 'ribosome biogenesis', 
  title = "cKO_Y15 vs cKO(NES =1.09996)",
  color = "red",
  subplots = 1:2,
  pvalue_table = TRUE
)

p3[[1]] <- p3[[1]] + 
  geom_hline(yintercept = 0, linetype = "solid", color = "grey50", linewidth = 0.5)


combined_plot <- wrap_elements(p1) / wrap_elements(p2) / wrap_elements(p3)

pdf('/data/ccz/project/mycn_seu/revise_data/revised_fig/gsea_ribosome.pdf', width = 4.5, height = 10)
print(combined_plot)
dev.off() 
```

```{r}
library(ggplot2)
library(patchwork)

p1 <- enrichplot::gseaplot2(
  specific_gsea_results$KY_vs_WT,
  geneSetID = 'ribosome biogenesis', 
  title = "cKO_Y15 vs Ctrl",
  color = "red",
  pvalue_table = TRUE
)

p1[[1]] <- p1[[1]] + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50", size = 0.5) +
  annotate("text", x = -Inf, y = Inf, hjust = -0.2, vjust = 2, 
           label = "NES = -2.05", size = 4.5, fontface = "bold")

p1[[3]] <- p1[[3]] + 
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50", size = 0.5)

combined_plot <- wrap_elements(p1) 

pdf('gsea_ribosome.pdf', width = 5, height = 12)
print(combined_plot)
dev.off()


save(specific_gsea_results,file = './revised_fig/gse_ribosome_result.RData')
```



